Appendix: Trends in age-adjusted mortality rates for blacks and whites by cause and sex, 1969-2013, United States

### 2017-02-21 Code/plot_age_standardardized_mortality_trends.Rmd
cleaning up

```{r load_libraries, echo=FALSE, warning=F, message=F}
library(dplyr)
library(purrr)
library(ggplot2)
library(stringr)
library(reshape2)
library(grid)
library(gridExtra)
library(plotly)
```

```{r}
included.states <- c("Alabama", "Arizona", "Arkansas", "California", "Colorado", "Connecticut",
"Delaware", "Florida", "Georgia", "Illinois", "Indiana", "Iowa", "Kansas", "Kentucky",
"Louisiana", "Maryland", "Massachusetts", "Michigan", "Minnesota", "Mississippi",   
"Missouri", "Nebraska", "Nevada", "New Jersey", "New Mexico", "New York",
"North Carolina", "Ohio", "Oklahoma", "Oregon", "Pennsylvania", "Rhode Island",
"South Carolina",  "Tennessee", "Texas", "Virginia", "Washington", "Washington DC",
"West Virginia", "Wisconsin")

completed.states <- included.states[17:24]
```

```{r, echo=F}

mortality.rates <- read.csv(paste0("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results/mortality_rates_", completed.states[1], ".csv"))
mortality.rates.diff <- read.csv(paste0("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results/mortality_rates_diff_", completed.states[1], ".csv"))

for(state_i in 1:length(completed.states)){
  mortality.rates <- rbind(mortality.rates, read.csv(paste0("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results/mortality_rates_", completed.states[state_i], ".csv")))

    mortality.rates.diff <- rbind(mortality.rates.diff, read.csv(paste0("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results/mortality_rates_diff_", completed.states[state_i], ".csv")))
}

 mortality.rates <- mortality.rates %>% mutate(state.reorder = reorder(state, as.numeric(Census_Division)))
 mortality.rates.diff <- mortality.rates.diff %>% mutate(state.reorder = reorder(state, as.numeric(Census_Division)))
# 
#  mortality.rates$state.reorder2 <- factor(mortality.rates$state, 
#                           levels = c("e1", "e2", "e3", "e4", "e5", "e6", "e7", "e8", "e9", "e10", "Maine",
#                                      "e11", "e12", "e13", "e14", "e15", "e16", "e17", "e18", "e19", "Vermont", "New Hampshire",
#                                      "Washington", "Idaho", "Montana", "North Dakota", "Minnesota", "Illinois", "Wisconsin", "Michigan", "New York", "Massachusettes", "Rhode Island",
#                                      "Oregon", "Nevada", "Wyoming", "South Dakota", "Iowa", "Indiana", "Ohio", "Pennsylvania", "New Jersey", "Connecticut", "e20",
#                                      "California", "Utah", "Colorado", "Nebraska", "Missouri", "Kentucky", "West Virginia", "Virginia", "Maryland", "Washington DC", "e21",
#                                      "e22", "Arizona", "New Mexico", "Kansas", "Arkansas", "Tennessee", "North Carolina", "South Carolina", "Delaware", "e23", "e24",
#                                      "e25", "e26", "e27", "Oklahoma", "Louisiana", "Mississippi", "Alabama", "Georgia", "e28", "e29",
#                                      "e30", "e31", "e32", "e33", "Texas", "e34", "e35", "e36", "e37", "Florida"))   
#     
#  data.sub <- subset(mortality.rates, sex == "Male" & COD == "Injuries")
#  test <- ggplot(data.sub, 
#                 aes(y = age.std.mortality.rate_med, x = year)) + 
#       geom_vline(aes(xintercept = 1983), col = "#636363", lty = 3) +
#       geom_vline(aes(xintercept = 1993), col = "#636363", lty = 3) +
#       geom_ribbon(aes(ymin = age.std.mortality.rate_LCL, ymax = age.std.mortality.rate_UCL, group = race, fill = Census_Division), col = NA, alpha = 0.5) +
#       geom_line(aes(col = Census_Division, lty = race)) +
#       facet_wrap( ~ state.reorder2, ncol = 11, drop = F) +
#       ylab("Age-standardized mortality rate (per 100,000)") +
#       xlab("Year") #+ 
#      # ggtitle(paste0("Age-adjusted mortality (per 100,000) in ", sex_i, " for ", cod))# + theme_minimal()
#  test

>>>>

```{r load_libraries, echo=FALSE, warning=F, message=F}
library(dplyr)
library(purrr)
library(ggplot2)
library(stringr)
library(reshape2)
library(grid)
library(gridExtra)
library(plotly)
library(forcats)

mortality.rates <- read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results/mortality_rates_combined.csv")
mortality.rates.diff <- read.csv("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Results/mortality_rates_diff_combined.csv")

mortality.rates <- mortality.rates %>% mutate(state.reorder = reorder(state, as.numeric(Census_Division)))
mortality.rates.diff <- mortality.rates.diff %>% mutate(state.reorder = reorder(state, as.numeric(Census_Division)))
```

```{r, echo=F}
# 
#  mortality.rates$state.reorder2 <- factor(mortality.rates$state, 
#                           levels = c("e1", "e2", "e3", "e4", "e5", "e6", "e7", "e8", "e9", "e10", "Maine",
#                                      "e11", "e12", "e13", "e14", "e15", "e16", "e17", "e18", "e19", "Vermont", "New Hampshire",
#                                      "Washington", "Idaho", "Montana", "North Dakota", "Minnesota", "Illinois", "Wisconsin", "Michigan", "New York", "Massachusettes", "Rhode Island",
#                                      "Oregon", "Nevada", "Wyoming", "South Dakota", "Iowa", "Indiana", "Ohio", "Pennsylvania", "New Jersey", "Connecticut", "e20",
#                                      "California", "Utah", "Colorado", "Nebraska", "Missouri", "Kentucky", "West Virginia", "Virginia", "Maryland", "Washington DC", "e21",
#                                      "e22", "Arizona", "New Mexico", "Kansas", "Arkansas", "Tennessee", "North Carolina", "South Carolina", "Delaware", "e23", "e24",
#                                      "e25", "e26", "e27", "Oklahoma", "Louisiana", "Mississippi", "Alabama", "Georgia", "e28", "e29",
#                                      "e30", "e31", "e32", "e33", "Texas", "e34", "e35", "e36", "e37", "Florida"))   
#     
#  data.sub <- subset(mortality.rates, sex == "Male" & COD == "Injuries")
#  test <- ggplot(data.sub, 
#                 aes(y = age.std.mortality.rate_med, x = year)) + 
#       geom_vline(aes(xintercept = 1983), col = "#636363", lty = 3) +
#       geom_vline(aes(xintercept = 1993), col = "#636363", lty = 3) +
#       geom_ribbon(aes(ymin = age.std.mortality.rate_LCL, ymax = age.std.mortality.rate_UCL, group = race, fill = Census_Division), col = NA, alpha = 0.5) +
#       geom_line(aes(col = Census_Division, lty = race)) +
#       facet_wrap( ~ state.reorder2, ncol = 11, drop = F) +
#       ylab("Age-standardized mortality rate (per 100,000)") +
#       xlab("Year") #+ 
#      # ggtitle(paste0("Age-adjusted mortality (per 100,000) in ", sex_i, " for ", cod))# + theme_minimal()
#  test
